---
title: "Clustering of antibody trajectories using abtrac"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{abtrac_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(abtrac)
```


## Example trajectories

Example trajectories are provided in `abtrac::trajectories`.

```{r}
trajectories <- abtrac::trajectories
```

`trajectory`: One antibody measured in one sample.
`t`: Timepoint, centered on the timepoint of interest (0), e.g., seroconversion.
`fc`: Fold change relative to baseline, `t==-1`.

```{r}
head(trajectories)
```


## Distance matrix

The distance matrix is computed using `ab_tr_dist()`.

```{r}
dist_mat <- ab_tr_dist(trajectories = trajectories)
```


## Clustering

Clustering is performed using the distance matrix. The range of desired number of clusters `ks` is selected. In the companion paper, the options `variant = "faster"` and `nstart = 10` were forwarded to `cluster::pam()`.

```{r}
pam_objects <- ab_tr_clust(dist_mat = dist_mat, ks = 4:10, verbose = FALSE, variant = "faster", nstart = 10)
```


```{r}
head(pam_objects)
```


## Selection of k - silhouette plot

The optimal cluster number `k` is determined using the silhouette method (here 7).

```{r, fig.width=7}
silhouette_plot(pam_objects = pam_objects)
```


## Model selection

Clusters are retrieved from the `pam` object and added to the original tibble of `trajectories`.

```{r}
trajectories_clust <- extract_clusters(trajectories = trajectories, pam_objects = pam_objects, sel_k = 7)
```


```{r}
head(trajectories_clust)
```


## Cluster visualization

The clusters are visualized using lineplots. Here, cluster 2 can be described as "delayed new-onset antibodies", cluster 5 as "stable new-onset antibodies", and cluster 7 as "transient new-onset antibodies".

```{r, fig.width=7, fig.height=7}
cluster_lineplot(trajectories_clust, ncol = 4)
```

## Notes

The expected install and run time for the package and vignette is less than 10 minutes on a "normal" laptop.
